---
title: "Introduction to Statistical Learning"
subtitle: "STAT 387" 
author: "Sang Xing"
format: 
  html:
    df-print: paged
    theme:
      light: materia
      dark: slate
    toc: true
    toc-title: Table of Contents
    toc-depth: 4
    toc-location: right
    number-sections: false
    number-depth: 3
    anchor-sections: true
    smooth-scroll: true
    link-external-icon: false
    link-external-newwindow: true
    code-fold: false
    code-overflow: scroll
    code-tools: 
      source: true
      toggle: true
      caption: none
    code-summary: "Show the code"
    highlight-style: atom-one
    link-external-filter: '^(?:http:|https:)\/\/www\.quarto\.org\/custom'
    html-math-method:
      method: mathjax
      url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
---

\vspace{-1cm}

```{r setup, include=FALSE}
library(tidyverse)  # Load core packages: 
# ggplot2,   for data visualization.
# dplyr,     for data manipulation.
# tidyr,     for data tidying.
# purrr,     for functional programming.
# tibble,    for tibbles, a modern re-imagining of data frames.
# stringr,   for strings.
# forcats,   for factors.
# lubridate, for date/times.
# readr,     for reading .csv, .tsv, and .fwf files.
# readxl,    for reading .xls, and .xlxs files.
# feather,   for sharing with Python and other languages.
# haven,     for SPSS, SAS and Stata files.
# httr,      for web apis.
# jsonlite   for JSON.
# rvest,     for web scraping.
# xml2,      for XML.
# modelr,    for modelling within a pipeline
# broom,     for turning models into tidy data
# hms,       for times.

library(magrittr)   # Pipeline operator
library(lobstr)     # Visualizing abstract syntax trees, stack trees, and object sizes
library(pander)     # Exporting/converting complex pandoc documents, EX: df to Pandoc table
library(ggforce)    # More plot functions on top of ggplot2
library(ggpubr)     # Automatically add p-values and significance levels  plots. 
# Arrange and annotate multiple plots on the same page. 
# Change graphical parameters such as colors and labels.
library(sf)         # Geo-spatial vector manipulation: points, lines, polygons
library(kableExtra) # Generate 90 % of complex/advanced/self-customized/beautiful tables
library(latex2exp)  # Latex axis titles in ggplot2
library(ellipse)    # Simultaneous confidence interval region to check C.I. of 2 slope parameters
library(plotly)     # User interactive plots
library(ellipse)    # Simultaneous confidence interval region to check C.I. of 2 regressors
library(olsrr)      # Model selections 
library(leaps)      # Regression subsetting 
library(pls)        # Partial Least squares
library(MASS)       # LDA, QDA, OLS, Ridge Regression, Box-Cox, stepAIC, etc,.
library(e1071)      # Naive Bayesian Classfier,SVM, GKNN, ICA, LCA
library(class)      # KNN, SOM, LVQ
library(ROCR)       # Precision/Recall/Sensitivity/Specificity performance plot 
library(boot)       # LOOCV, Bootstrap,
library(caret)      # Classification/Regression Training, run ?caret::trainControl
library(corrgram)   # for correlation matrix
library(corrplot)   # for graphical display of correlation matrix

set.seed(1234)        # make random results reproducible

current_dir <- getwd()

if (!is.null(current_dir)) {
  setwd(current_dir)
  remove(current_dir)
}
```

```{css, echo=FALSE}
body {
font-family: "Times New Roman";
font-size: 0.8em;
}

h2, .h2, h2.anchored, 
h3, .h3, h3.anchored,
.quarto-title, 
.quarto-title-meta {
text-align: center;
}

h2, .h2 {
font-size: 1.5em;
}

h3, .h3 {
font-size: 1.25em;
}

h4, .h4 {
font-size: 1em;
}

h5, .h5 {
font-size: 0.75em;
}

h6, .h6 {
font-size: 0.7em;
}

.quarto-title-meta {
display: block!important;
}

.quarto-title-meta-heading {
display: none;
}

code {
color: #61AFEF;
}

.btn {
border-color: #f8f9fa9e!important;
}

body.quarto-dark {
background-color: #282c34;

color: #d8dee9;

.pagedtable .even {
background-color: #3b4252!important;
}

p code:not(.sourceCode), li code:not(.sourceCode) {
background-color: #4c566a!important;
}
}
```


## Preamble

Consider the wine quality dataset from [UCI Machine Learning Respository](https://archive.ics.uci.edu/ml/datasets/Wine+Quality). We will focus only on the data concerning white wines (and not red wines). Dichotomize the `quality` variable as `good`, which takes the value 1 if quality ≥ 7 and the value 0, otherwise. We will take `good` as response and all the 11 physiochemical characteristics of the wines in the data as predictors.

### Problem Statements

Use 10-fold cross-validation for estimating the test error rates below and compute the estimates using `caret` package with seed set to 1234 before each computation.

(a) Fit a KNN with K chosen optimally using test error rate. Report error rate, sensitivity, specificity, and AUC for the optimal KNN based on the training data. Also, report its estimated test error rate.
(b) Repeat (a) using logistic regression.
(c) Repeat (a) using LDA.
(d) Repeat (a) using QDA.
(e) Compare the results in (a)-(d). Which classifier would you recommend? Justify your answer.



### Methodologies
- KNN
- GLM/logit/glmnet
- LDA
- QDA
- Naive Bayes
- Random Forest
- Decision Tree
- Regression Tree
- SVM
- Regression/Multi-Regression


## Exploratory Analysis


### Data Import
```{r data_import}
wine.data <- read.csv("data\\winequality-white.csv", sep=";", header = T)

wine.data <-  wine.data %>% mutate(good = ifelse(quality>=7, 1, 0))

head(wine.data)
```

### Basic Analysis
```{r basic_analysis}
dim(wine.data)

#Removing the Duplicate Rows
wine.data <- wine.data[!duplicated(wine.data), ]

dim(wine.data)

str(wine.data)
summary(wine.data)

#Check for NA in dataset
sum(is.na(wine.data))

#Response count in the dataset
table(wine.data$quality)
```

### Data Histograms
```{r hist, warning=FALSE}
wine.colnames <- colnames(wine.data[, 1:12])
num_plots     <- length(wine.colnames)
num_rows      <- ceiling(num_plots/3)


# Create an empty list to store plots
grid_arr      <- list()


# Loop over each column name in the wine.colnames vector
for(i in 1:num_plots) {
  # Create a ggplot object for the current column using aes
  plt <- ggplot(data = wine.data, aes_string(x = wine.colnames[i])) +
    geom_histogram(binwidth = diff(range(wine.data[[wine.colnames[i]]]))/30, 
                   color = "black", fill = "slategray3") +
    labs(x = wine.colnames[i], y = "Frequency") +
    theme_bw()
  
  # Add the current plot to the grid_arr list
  grid_arr[[i]] <- plt
}

grid_arr <- do.call(gridExtra::grid.arrange, c(grid_arr, ncol = 3))

remove(grid_arr)
remove(plt)
```

### Data Relationships
```{r relation_plot, fig.width=10, fig.height=5, message=FALSE}
reshape2::melt(wine.data[, 1:12], "quality") %>% 
  ggplot(aes(value, quality, color = variable)) +  
  geom_point() + 
  geom_smooth(aes(value,quality, colour=variable), method=lm, se=FALSE)+
  facet_wrap(.~variable, scales = "free")

#Collinearity between Attributes
cor(wine.data) %>% 
  corrplot::corrplot(method = 'number',  type = "lower", tl.col = "steelblue", number.cex = 0.5)
```


## K-Nearest Neightbor



## Logistic Regression


## Linear Discriminant Analysis


## Quadratic discriminant analysis

